### 2021-01-07 results/notes.tex

\documentclass[prd,amsmath,amssymb,floatfix,superscriptaddress,nofootinbib]{revtex4-1}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{epsfig}
\usepackage{color}
\usepackage{natbib}
\usepackage{textcase}
\usepackage{graphicx}
\usepackage{ifthen}
\usepackage{xstring}
\usepackage{graphicx}
\usepackage[utf8]{inputenc} 
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{epstopdf}
\epstopdfsetup{update}
\DeclareGraphicsExtensions{.ps, .png}
\epstopdfDeclareGraphicsRule{.ps}{pdf}{.pdf}{ps2pdf -dEPSCrop -dNOSAFER #1 \OutputFile} 
\usepackage{dcolumn} 
\usepackage{multirow}
\usepackage{appendix}
\usepackage{footnote}
\usepackage{tabularx,ragged2e,booktabs}
\usepackage[normalem]{ulem}
\usepackage{float}
\restylefloat{table}

\newcommand{\tauhi}{$\tau_{\rm hi}\,$}
\newcommand{\taulo}{$\tau_{\rm lo}\,$}

\newcommand{\refsec}[1]{section~\ref{sec:#1}}
\newcommand{\refeq}[1]{Eq.~(\ref{eq:#1})}
\newcommand{\refssec}[1]{section~\ref{subsec:#1}}
\newcommand{\reffig}[1]{Fig.~\ref{fig:#1}}
\newcommand{\refFig}[1]{Fig.~\ref{fig:#1}}
\newcommand{\curv}{{\cal R}}
\newcommand{\xef}{x_e^{\rm fid}}
\newcommand{\xmax}{x_e^{\rm max}}
\newcommand{\zmax}{z_{\rm max}}
\newcommand{\zmin}{z_{\rm min}}
\newcommand{\xemin}{x_e^{\rm min}}

\newcommand{\ra}{\rightarrow}
\def\max{_{\mathrm{max}}}
\def\lsim{\mathrel{\raise.3ex\hbox{$$<$$\kern-.75em\lower1ex\hbox{$\sim$}}}}
\def\gsim{\mathrel{\raise.3ex\hbox{$$>$$\kern-.75em\lower1ex\hbox{$\sim$}}}}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\newcommand{\wh}[1]{\textcolor{blue}{#1}}
\newcommand{\ch}[1]{\textcolor{red}{#1}}

\def\mnras{Mon.\ Not.\ R.\ Astron.\ Soc.\ }
\definecolor{darkgreen}{cmyk}{0.85,0.2,1.00,0.2} 
\definecolor{purple}{cmyk}{0.5,1.0,0,0} 
\def\physrep{Phys.~Rep.}

\definecolor{ultramarine}{rgb}{0.07, 0.04, 0.56}
\definecolor{cadmiumgreen}{rgb}{0.0, 0.42, 0.24}
\definecolor{indigo(dye)}{rgb}{0.0, 0.25, 0.42}
\usepackage[linktocpage=true]{hyperref}
\hypersetup{
colorlinks=true,
citecolor=ultramarine,
linkcolor=cadmiumgreen,
urlcolor=indigo(dye),
pdfauthor={},
pdftitle={},
pdfsubject={}
}

\begin{document}

\title{Reionization Planck 2018 Notes}

\author{Chen Heinrich}
\author{Wayne Hu}

\maketitle

\section{Introduction}
\label{sec:intro}

Trying to collect all of our tests here.\\

List of tests:\\
\begin{itemize}
    \item {Direct MCMC chains
        \begin{enumerate}
            \item {PC zmax = 30}
            \item {PC zmax = 50}
            \item{Tanh model}
            \item{Two-parameter model with \taulo and \tauhi}
        \end{enumerate}
    }
    \item{Best-fit tests
        \begin{enumerate}
            \item {From the chains for the 2-parameter model, we found that if we restrict \tauhi $> 0.02$, we get that the ML model is: 
            \begin{itemize}
                \item {\tauhi = 0.0206}
                \item {\taulo = 0.0414} 
                \item {$z_{\rm re}$ = 6.29} 
                \item {$x_{e, \mathrm{hi}}$ = 0.0748}
                \item {-loglike = 499.5458} 
            \end{itemize}
            There is a difference of $\Delta \chi^2 = 2.6$, compared against the ML model from the full chain:
            \begin{itemize}
                \item {\tauhi = 0.00542}
                \item {\taulo = 0.0543} 
                \item {$z_{\rm re}$ = 7.68} 
                \item {$x_{e, \mathrm{hi}}$ = 0.0214}
                \item {-loglike = 498.2425} 
            \end{itemize}
             }
            \item{Using above information, we ran a best-fit search fixing \tauhi = 0.02 to find the best-fit model:
            \begin{itemize}
                \item {\tauhi = 0.02 (fixed)}
                \item {\taulo = 0.045 (c.f. 0.0414 from chains)} 
                \item {$z_{\rm re}$ = 6.67} 
                \item {$x_{e, \mathrm{hi}}$ = 0.073}
                \item {-loglike = 499.2914} 
            \end{itemize} 
            giving a difference of $\Delta \chi^2 = 2.1$  (c.f. 2.6 above).}\\
            Note: I need to update the plots with xe(z) to this model.
            \item{We also ran a chain fixing \taulo = 0.04, but I forgot what this was for ...}
        \end{enumerate}
        }

    \item {KDE tests
        \begin{enumerate}
            \item {The tanh model: direct MCMC, vs python (f = 0.10, 0.12, 0.14) and fortran version (0.14) - coming up}
            \item {The 2-parameter model: direct MCMC vs fortran version (f = 0.10, 0.12, 0.14) - coming up}
        \end{enumerate}
    }

    \item {Preparation for manuscript
        \begin{enumerate}
            \item {Testing \taulo prior (\taulo $\geq 0.04$) - Figs~\ref{fig:taulo_prior_test_cl} and~\ref{fig:tau_lo_prior_test_xe}: $\chi^2 = 0.03$ for $l\in[2, 100]$ between $C_l^{EE}$ from full vs PC projected $x_e(z)$ for the model (\taulo, \tauhi) = (0.04, 0.02).}
            \item {Plots to remake for paper: 1) }
        \end{enumerate}
    }

\end{itemize}

\begin{figure}
\includegraphics[width=0.65\textwidth]{results/direct_mcmc/pl18_plots_zmax30/plot_pub_tau_gtz_dz_0p1_pl18_pc_zmax30_pliklite_post_0930_and_pl18_pc_zmax30_pliklite_srollv2_0930.pdf}
\caption{PC chains for zmax = 30. Planck 2018 original lowE vs srollv2 likelihood (plik\_lite\_TTTEEE + lowl + simall\_EE vs plik\_lite\_TTTEEE + lowl + sroll2\_EE).
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{results/direct_mcmc/pl18_plots_zmax50/plot_pub_tau_gtz_dz_0p1_pl18_pc_zmax50_pliklite_post_and_pl18_pc_zmax50_pliklite_srollv2.pdf}
\caption{PC chains for zmax = 50. Planck 2018 original lowE vs srollv2 likelihood (plik\_lite\_TTTEEE + lowl + simall\_EE vs plik\_lite\_TTTEEE + lowl + sroll2\_EE).
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{results/direct_mcmc/pl18_plots_zmax30/plot_pub_tau_gtz_dz_0p1_pl18_pc_zmax30_pliklite_post_0930_and_pl18_pc_zmax50_pliklite_post.pdf}
\caption{Comparing zmax = 30 and 50 PC chains using Planck 2018 original lowE likelihood (plik\_lite\_TTTEEE + lowl + simall\_EE). Note that the zmax = 30 uses 5 PCs, whereas the zmax = 50 chains uses 7 PCs.
}
\label{fig:tau_gtz_zmax_30_vs_50_simall_EE}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{results/direct_mcmc/pl18_plots_zmax30/plot_pub_tau_gtz_dz_0p1_pl18_pc_zmax30_pliklite_srollv2_0930_and_pl18_pc_zmax50_pliklite_srollv2.pdf}
\caption{Same as Fig.~\ref{fig:tau_gtz_zmax_30_vs_50_simall_EE}, but for the srollv2 likelihood (plik\_lite\_TTTEEE + lowl + sroll2\_EE).
}
\label{fig:}
\end{figure}

%Optional
\begin{figure}
\includegraphics[width=0.65\textwidth]{results/direct_mcmc/pl18_plots_zmax30/plot_pub_tau_gtz_dz_0p1_pl18_pc_zmax30_plikfull_and_pl18_tanh_post_plikfull.pdf}
\caption{PC zmax = 30 vs tanh chains with plik\_full\_TTTEEE for the high-l likelihood.
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{results/direct_mcmc/two_parameter_model/tauhi_taulo_chains/pl18_tanh_highz_test2_run1_tri.png}
\caption{Direct MCMC chains of the two-parameter model with \tauhi and \taulo: triangle plot of \tauhi and \taulo. The marginalized 1D constraints are \taulo = $0.0516 \pm 0.0076$, \tauhi = $0.0100 \pm   0.0066$. The ML model is \taulo = 0.0543 and \tauhi = 0.0054 corresponding to $z_{\rm re} = 7.68$ and $x_{e, \mathrm{min}} = 0.021$.
}
\label{fig:two_parameter_model_2D_plot}
\end{figure}

 \begin{figure}
\includegraphics[width=0.65\textwidth]{   results/direct_mcmc/pl18_plots_zmax30/plot_pub_tau_gtz_dz_0p1_pl18_pc_zmax30_pliklite_srollv2_0930_and_pl18_tanh_post_pliklite_srollv2_with_added_two_parameter_ML_xe_full.pdf}
\caption{PC zmax = 30 vs tanh chains with plik lite and srollv2, with ML model from two-parameter model for \tauhi restricted to $>$ 0.02 (ML model: \tauhi = 0.0206, \taulo = 0.0414). (Note: I need to update the xe(z) curve to be that of the best-fit model in the best-fit search with tau_hi = 0.02 fixed.)
}
\label{fig:two_parameter_model_ML}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{  results/direct_mcmc/pl18_plots_zmax30/plot_pub_tau_gtz_dz_0p1_pl18_pc_zmax30_pliklite_srollv2_0930_and_pl18_tanh_post_pliklite_srollv2_with_added_two_parameter_ML_PC_proj.pdf}
\caption{Same as Fig.~\ref{fig:two_parameter_model_ML} except the blue curve is a PC projection instead of the full xe(z) of the model.
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{   results/direct_mcmc/two_parameter_model/fixed_taulo_chains/plot_1D_tau_hi_pl18_tanh_highz_fixed_taulo_test2_run1.pdf}
\caption{Chains for two-parameter model with fixed \taulo = 0.04: 1D distribution of \tauhi (not sure what this is for anymore...).}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{python_kde/pl18_pc_zmax30_pliklite_srollv2_1015_tau_posterior_fraccov_1p0_burnin_10000_yes_norm_gaussian0p1_0p12_0p14.pdf}
\caption{KDE tests in python for the tanh model with various f = 0.1, 0.12, 0.14. It is probably safest to use any of these, so given the smaller f the better, f = 0.1 is probably the best choice here.
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/direct_mcmc_vs_python_kde_and_cosmomc_kde_tanh_tau.png}
\caption{LATEST (to be debugged) -- Added cosmomc KDE version of tau posterior computed with f = 0.14, against python versions with f = 0.1, 0.12, 0.14. They do not agree ... (Note: need to make sure python version had the same covariance matrix, and length of chain ... should probably rerun python version actually.)
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/pl18_relike_tanh_tau_dz0p5_1D_direct_vs_fortran_and_python_kde.png}
\caption{DEBUGGING - updated the KDE tanh results to have tanh dz = 0.5. Also added in faint blue lines Gaussian and convolved Gaussian with $\sigma \rightarrow \sigma \sqrt(1+f)$.
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/pl18_relike_tanh_tau_dz0p5_1D_direct_vs_fortran_and_python_kde_double_checked_python.png}
\caption{DEBUGGING - updated the python curves with the same covariance matrix, chains, etc (also confirmed that tanh dz = 0.5). 
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{ cosmomc_kde/pl18_relike_tanh_tau_dz0p5_1D_direct_vs_fortran_kde.png}
\caption{DEBUGGING - tanh dz = 0.5 for all curves. Same as above, but only keeping some curves for clarity.)
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/pl18_relike_tanh_tau_dz0p5_1D_direct_vs_fortran_and_python_kde_with_grid_test.png}
\caption{NEW - added blue circles which are evaluating the Fortran KDE code on a grid of 100 values of $\tau$. This shows that the extra smoothing in the red curve (Fotran MCMC code) vs green (python grid evaluations) are \textbf{likely due to the smoothing used in getdist to make the red curve}.
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/pl18_relike_tanh_tau_dz0p5_1D_direct_vs_fortran_vs_mock_gaussian_fp10_fp14.png}
\caption{NEW - comparing f = 0.10 (orange dotted) and 0.14 (green dotted) in python KDE (latest version) against an example of gaussian smoothing with f = 0.10 (orange dashed) and 0.14 (blue dashed) on a Gaussian function (blue solid).
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/pl18_tanh_highz_test2_run1_vs_relike_tanh_highz_test1_run1_f0p14_tri.png}
\caption{version 1 (has bug that \taulo is not limited to where PCs can model z>6): 2-parameter model 2D plot from cosmomc KDE with f = 0.14 vs direct MCMC.
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/pl18_tanh_highz_test2_run1_vs_relike_tanh_highz_test1_run1_f0p14_taulo_prior_0p0415_tri.png}
\caption{version 2 (added prior \taulo $\leq 0.0415$ for the KDE version, f = 0.14). Direct MCMC chain with the same prior is still running.
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/pl18_tanh_highz_test2_run1_vs_relike_tanh_highz_test1_run1_f0p14_and_f0p20_taulo_prior_0p0415_tri.png}
\caption{version 3, added a curve with f = 0.20 (blue).
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/pl18_tanh_highz_test2_run1_vs_relike_tanh_highz_test1_run1_f0p14_and_f0p20_taulo_prior_0p0415_tri.png}
\label{fig:}
\caption{NEW: Direct MCMC without prior (gray) vs with prior (red) vs KDE fortran with same prior (\taulo \geq 0.0415) and $f = 0.14$.
} 
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/getdist_smoothing_test/pl18_relike_tanh_tau_dz0p5_1D_gauss_smoothed_tanh_std_0p0067.png}
\label{fig:}
\caption{Gaussian convolving the direct MCMC 1D distribution by hand with f = 0.14 where the original standard deviation is set to 0.0067. This corresponds to the right side of asymmetric distribution $\tau = 0.0591 + 0.0055 - 0.0067$ from the original MCMC.
} 
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/getdist_smoothing_test/pl18_relike_tanh_tau_dz0p5_1D_gauss_smoothed_tanh_std_0p0055.png}
\label{fig:}
\caption{Gaussian convolving the direct MCMC 1D distribution by hand with f = 0.14 where the original standard deviation is set to 0.0055. This corresponds to the left side of asymmetric distribution $\tau = 0.0591 + 0.0055 - 0.0067$ from the original MCMC.
} 
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/getdist_smoothing_test/pl18_pc_zmax30_pliklite_srollv2_1015_tau_posterior_python_kde_f_0p1_vs_0p14.png}
\label{fig:}
\caption{Comparing f = 0.10 and f = 0.14 lines from python KDE - to see that f = 0.14 is indeed more smoothed than f = 0.10 as expected.
} 
\end{figure}

\begin{figure}
\includegraphics[width=0.4\textwidth]{cosmomc_kde/getdist_smoothing_test/pl18_relike_tanh_tau_dz0p5_1D_smooth_scale_1D_2_num_bins_1D_1600.png}
\includegraphics[width=0.4\textwidth]{cosmomc_kde/getdist_smoothing_test/pl18_relike_tanh_tau_dz0p5_1D_smooth_scale_1D_2_num_bins_1D_800.png}
\includegraphics[width=0.4\textwidth]{cosmomc_kde/getdist_smoothing_test/pl18_relike_tanh_tau_dz0p5_1D_smooth_scale_1D_2_num_bins_1D_100.png}
\includegraphics[width=0.4\textwidth]{cosmomc_kde/getdist_smoothing_test/pl18_relike_tanh_tau_dz0p5_1D_smooth_scale_1D_2_num_bins_1D_50.png}
\label{fig:}
\caption{Changing num\_bins = 1600 for the Fortran/cosmomc KDE w/ f = 0.14, which recovers close to the python KDE line w/ the same f. Lowering num\_bins increases smoothing.
} 
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/taulo_prior_test/plot_cls_taulo_0p04_tauhi_0p02.png}
\caption{Testing \taulo prior -- testing that \taulo $\geq$ 0.04 as \taulo prior is OK, by calculating $\chi^2$ between $C_l$ from full $x_e(z)$ vs PC decomposition of the model (\taulo, \tauhi) = (0.04, 0.02). The $\chi^2 = \sum_{2}^{l_{\rm max}} (\Delta C_l^{EE})^2 / \mathrm{Cov}_l = 0.03$ where $\mathrm{Cov}_l = 2 (C_l^{EE})^2/(2l+1)$. 
} 
\label{fig:taulo_prior_test_cl}
\end{figure}

\begin{figure}
\includegraphics[width=0.65\textwidth]{cosmomc_kde/taulo_prior_test/plot_xez_taulo_0p04_tauhi_0p02.png}
\caption{Testing \taulo prior -- Full $x_e(z)$ of the model (\taulo, \tauhi) = (0.04, 0.02) shown in the Fig.~\ref{fig:taulo_prior_test_cl}. 
}
\label{fig:tau_lo_prior_test_xe}
\end{figure}

\begin{widetext}

\begin{figure}
\includegraphics[width=0.7\textwidth]{results/pc_results/plot_mj_triangle_pl18_pc_zmax30_pliklite_srollv2_vs_pl18_pc_zmax30_pliklite.pdf}
\caption{PC chains for zmax = 30 using Planck 2018 srollv2 likelihood vs Planck 2015.
}
\label{fig:plot_mjs}
\end{figure}

\begin{figure}
\includegraphics[width=0.4\textwidth]{results/pc_results/plot_mj_triangle_t18_r12_t19_t20_vs_pl18_pc_zmax30_pliklite.pdf}
\caption{PC chains plane for zmax = 30 using Planck 2018 srollv2 likelihood vs Planck 2015.
}
\label{fig:plot_mjs_2018_vs_2015}
\end{figure}

\end{widetext}

\begin{figure}
\includegraphics[width=0.4\textwidth]{results/pc_results/plot_m1_m2_pl18_pc_zmax30_pliklite_srollv2_vs_pl18_pc_zmax30_pliklite_wTauTrajectory.pdf}
\caption{m1-m2 plane for zmax = 30 using Planck 2018 original lowE vs srollv2 likelihood.
}
\label{fig:plot_m1m2}
\end{figure}

Tests done for tanh high-z model:

Problem:
1) For the smaller tanh width at low z, now set to be dz = 0.015(1+z), the model with parameters taulo = 4.0, tauhi = 0.0 gives Cl = NaN while applying smoothing or using tauhi = 1e-4 gives sensible Cls. For the original fixed dz = 0.25, this was not problem.

Observations and solution:
1) It is automatically set that xemin = 0.0 if tauhi < some threshold. 
2) xemin = 0.0 gives Cl = NaN because of precision issues.
3) In fact, xemin < approximately 1e-18 would also give Cl = NaN.
4) Instead of setting xemin to 0.0 for tauhi below some threshold, we let iterative search converge by itself.

Additional fixes:
1) We set xemin to sign(xemin) * 1e-7 if |xemin| < 1e-7. 
2) We start the xemin search between -0.1 and 100 instead of 0.0 and 100, to allow for negative values of xemin to be searched. We checked that setting xemin = -1e-18 does not cause a problem.
3) Smoothing capability is added to smooth xe(z), but this is not used for it is too slow.

The above changes are implemented in the branch chen/debug\_cl\_nan.

-----

\begin{figure}
\includegraphics[width=0.4\textwidth]{results/cosmomc_kde/taulo_prior_test/relike_tanh_highz_f0p14_taulo_prior_0p04_with_derived_parameters_zre_greater_than_6p25_tri.png}
\caption{Illustration of the bug in the tanh high-z model, introduced before running chains w/ taulo prior 0.04 instead of 0.0415.
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.4\textwidth]{cosmomc_kde/plot_tanh_highz_direct_mcmc_with_taulo_priors.png}
\caption{Previous plot of direct MCMC w/ various \taulo priors showing that the MCMC runs look fine and inferring bug is on the KDE side of the code.
} 
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.4\textwidth]{results/cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test8_run9_f0p14_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri.png}
\caption{After bug fix: direct MCMC ($z_{\rm re} > 6.1$) vs KDE ($z_{\rm re} > 6.1$) done by hand) everything is in good agreement! This is applying the bug fix on the latest code, i.e. w/ new definition of tanh high-z model that adopts a smaller $dz = 0.015 (1+z)$ at the low-z transition and boundary point for \taulo and \tauhi defined at $z_{\rm re} + 1 \times dz$ instead of $z_{\rm re} + 2 \times dz$ and various other minor changes (e.g. $|x_{e,\rm min}|<10^{-7}$ is sent to $x_{e,\rm min} = \mathrm{sgn}(x_{e,\rm min}) \times 10^{-7}$, $\tau$ convergence change, etc.) —  
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.4\textwidth]{results/cosmomc_kde/getdist_smoothing_test/pl18_tanh_highz_test3_run1_vs_relike_tanh_highz_test1_run1_f0p10_taulo_prior_0p0415_0p10_0p14_0p20_tri_from_getdist_chains_smooth_scale_1D_-1_num_bins_1D_100.png}
\caption{Previous plot of direct MCMC vs KDE for the old tanh high-z model (less sharp transition, etc) for various f = 0.10, 0.14, 0.20. (CH: Same test to be done for the after bug fixed version shown in the previous plot.)
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.4\textwidth]{cosmomc_kde/pl18_relike_tanh_tau_dz0p5_oversampled_1D_smooth_scale_1D_2_num_bins_1D_1600_post_bug_fix.png}
\includegraphics[width=0.4\textwidth]{cosmomc_kde/pl18_relike_tanh_tau_dz0p5_oversampled_1D_smooth_scale_1D_2_num_bins_1D_800_post_bug_fix.png}
\includegraphics[width=0.4\textwidth]{cosmomc_kde/pl18_relike_tanh_tau_dz0p5_oversampled_1D_smooth_scale_1D_2_num_bins_1D_100_post_bug_fix.png}
\includegraphics[width=0.4\textwidth]{cosmomc_kde/pl18_relike_tanh_tau_dz0p5_oversampled_1D_smooth_scale_1D_2_num_bins_1D_50_post_bug_fix.png}
\label{fig:}
\caption{Oversampled tanh tau posterior after fixing bug and with even more samples than previously. Changing num\_bins = 1600, 800, 100, 50 for the Fortran/cosmomc KDE w/ f = 0.14, which entirely recovers (for num\_bins = 100) close to the python KDE line w/ the same f which is computed as a an evaluation of the effective likelihood at 100 $\tau$ values rather than MCMC sampling. This shows that the discrepancy seen previously between the two was only due to smoothing applied by getdist on the MCMC samples when very few samples were present in the chains.
} 
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=1.0\textwidth]{paper/plots/plot_mj_triangle_t18_r12_t19_t20_vs_pl18_pc_zmax30_pliklite_srollv2_1015_wTauTrajectory_pl15_wTanhML_wGaussEllipse.pdf}
    \caption{Caption}
    \label{fig:plot_triangle_wGauss}
\end{figure}

\bibliography{rei.bib}

\begin{figure}
\includegraphics[width=0.7\textwidth]{ cosmomc_kde/pl18_relike_tanh_tau_dz0p5_oversampled_1D_smooth_scale_1D_2_num_bins_1D_100_post_bug_fix_wGauss.png}
\caption{Same as previous figure for num\_bins = 100, but adding also the posterior obtained from the Gaussian likelihood using directly the covariance and mean of the PC chains.
}
\label{fig:}
\end{figure}

\begin{figure}
\includegraphics[width=0.4\textwidth]{cosmomc_kde/plot_ellipse.png}
\caption{Examining the variation in the ellipse area expected when a mock covariance is multiplied by 1.0, 1.14 and 1.28.
}
\label{fig:}
\end{figure}

\begin{figure}
%
\includegraphics[width=0.35\textwidth]{cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test9_run1_f0p28_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri.png}
\includegraphics[width=0.35\textwidth]{results/cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test9_run1_f0p28_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri_smooth_scale_1D_2_num_bins_1D_20_smooth_scale_2D_2_num_bins_2D_25.png}
%
\includegraphics[width=0.35\textwidth]{cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test9_run4_f0p18_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri.png}
\includegraphics[width=0.35\textwidth]{cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test9_run4_f0p18_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri_smooth_scale_1D_2_num_bins_1D_20_smooth_scale_2D_2_num_bins_2D_25.png}
%
\includegraphics[width=0.35\textwidth]{cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test8_run9_f0p14_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri.png}
\includegraphics[width=0.35\textwidth]{cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test8_run9_f0p14_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri_smooth_scale_1D_2_num_bins_1D_20_smooth_scale_2D_2_num_bins_2D_25.png}
%
\includegraphics[width=0.35\textwidth]{cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test9_run3_f0p10_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri.png}
\includegraphics[width=0.35\textwidth]{cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test9_run3_f0p10_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri_smooth_scale_1D_2_num_bins_1D_20_smooth_scale_2D_2_num_bins_2D_25.png}
%
\label{fig:}
\caption{Left: Automated getdist smoothing; Right: hand tuned smoothing (slightly aggressive binning).
} 
\end{figure}

%\begin{figure}
%    \centering
%    \includegraphics{cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test8_run9_f0p14_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0393_tri.png}
%    \caption{$\taulo$ prior imposed at 0.0393, giving a straight cut in the 2D constraint, rather than the rounded zre = 6.1 prior.}
%    \label{fig:my_label}
%\end{figure}

\begin{figure}
%
\includegraphics[width=0.35\textwidth]{results/cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test9_run1_f0p28_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri_smooth_scale_1D_0.3_num_bins_1D_100_smooth_scale_2D_0.3_num_bins_2D_100.png}
\includegraphics[width=0.35\textwidth]{results/cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test9_run1_f0p28_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri_smooth_scale_1D_0.5_num_bins_1D_100_smooth_scale_2D_0.5_num_bins_2D_100.png}
%
\includegraphics[width=0.35\textwidth]{results/cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test9_run4_f0p18_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri_smooth_scale_1D_0.3_num_bins_1D_100_smooth_scale_2D_0.3_num_bins_2D_100.png}
\includegraphics[width=0.35\textwidth]{results/cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test9_run4_f0p18_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri_smooth_scale_1D_0.5_num_bins_1D_100_smooth_scale_2D_0.5_num_bins_2D_100.png}
%
\includegraphics[width=0.35\textwidth]{results/cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test8_run9_f0p14_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri_smooth_scale_1D_0.3_num_bins_1D_100_smooth_scale_2D_0.3_num_bins_2D_100.png}
\includegraphics[width=0.35\textwidth]{results/cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test8_run9_f0p14_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri_smooth_scale_1D_0.5_num_bins_1D_100_smooth_scale_2D_0.5_num_bins_2D_100.png}
%
\includegraphics[width=0.35\textwidth]{results/cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test9_run3_f0p10_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri_smooth_scale_1D_0.3_num_bins_1D_100_smooth_scale_2D_0.3_num_bins_2D_100.png}
\includegraphics[width=0.35\textwidth]{results/cosmomc_kde/pl18_tanh_highz_test5_run1_vs_relike_tanh_highz_test9_run3_f0p10_taulo_prior_0p03_zre_prior_6p1_taulo_prior_0p0_tri_smooth_scale_1D_0.5_num_bins_1D_100_smooth_scale_2D_0.5_num_bins_2D_100.png}
%
\label{fig:}
\caption{Left: getdist smooth\_scale\_1D = smooth\_scale\_2D = 0.3; Right: same as left but with 0.5.
} 
\end{figure}

\end{document}